Characterizing subarctic European vegetation for land surface modeling of ozone pollution and climate risk

### 2021-09-27 VegOzone.tex
Rework section 2 intro

In our assessment of primary drivers of ozone damage risk on vegetation in the subarctic, we focus on conditions during the 2018/19 growing season at the NIBIO operated Environment Centre Svanhovd. We also set these into a climatological and regional context. To this end, we collected atmospheric and ozone monitoring data taken at Svanhovd and additional ozone monitoring data from observation sites in northern Finland and Sweden for cross-evaluation. Varying update frequencies and length of available data sets made it necessary to fetch the data from various sources (further details below). Svanhovd synonymously refers to the atmospheric monitoring site Svanvik/Pasvik throughout the paper.\\

@@

To relate local ozone concentration (\chem{[O_3]}) with visible damage on ozone sensitive plants, both an ozone bio-monitor (referred to as ozone garden) and a conventional ozone monitor using the existing atmospheric monitoring infrastructure had been installed during the 2018/19 growing season at the NIBIO centre. The locations of the atmospheric monitoring site and the ozone garden are marked in the aerial photography shown in Fig.~\ref{fig:svanhovd_research_station}a). An ozone garden consists of selected plant species that are sensitive to ozone and are likely to display visible injuries. Cultivated species were, e.g. clover, tobacco, and potatoes. As shown in Fig.~\ref{fig:svanhovd_research_station}b), the plants had to be protected from herbivores with a wire-mesh fence. In 2018, we qualitatively observed visible ozone damage on semi-natural vegetation (clover) and crops (tobacco). In contrast, no such visible damage was found on the clovers in 2019, although it was observed on the sensitive tobacco cultivar. This indicates that the vegetation may have been more affected by ozone in 2018 than in 2019.

All measured atmospheric variables including ozone for the 2018/19 growing season were obtained through \href{luftkvalitet.no}{luftkvalitet.no} operated by \citet{NILU_AIRQ}. The ozone monitoring data have been also added to the EBAS database operated by \citet{NILU_EBAS} and are thus openly accessible. All long-term ozone observation data described in Section~\ref{subsec:ozone_data} are also obtained from EBAS. The long-term data sets of agrometeorological variables including temperature and precipitation are available from September 1992 to the present day \citep[LandbruksMeteorologiske Tjeneste][note the station name here is Pasvik]{LMT_NIBIO}.

>>>>

In our assessment of primary drivers of ozone damage risk on vegetation in the subarctic, we focus on conditions during the 2018/19 growing season at the NIBIO operated Environment Centre Svanhovd. We also set these into a climatological context. To this end, we collected atmospheric and ozone monitoring data taken at Svanhovd over the past $35\,\unit{years}$. Svanhovd synonymously refers to the atmospheric monitoring site Svanvik/Pasvik throughout the paper.\\

@@

All measured atmospheric variables including ozone for the 2018/19 growing season were obtained through \href{luftkvalitet.no}{luftkvalitet.no} operated by \citet{NILU_AIRQ}. The ozone monitoring data have been also added to the EBAS database operated by \citet{NILU_EBAS} and are thus openly accessible. Long-term ozone observation data (1986--1996) were obtained from EBAS. The long-term data sets of agrometeorological variables including temperature and precipitation are available from September 1992 to the present day \citep[LandbruksMeteorologiske Tjeneste][note the station name here is Pasvik]{LMT_NIBIO}.

To relate local ozone concentration \chem{[O_3]} with visible damage on ozone sensitive plants, both an ozone bio-monitor (referred to as ozone garden) and a conventional ozone monitor using the existing atmospheric monitoring infrastructure had been installed during the 2018/19 growing season at the NIBIO centre. We will report \chem{[O_3]} as volume mixing ratios (VMR) in \unit{ppb}. The locations of the atmospheric monitoring site and the ozone garden are marked in the aerial photography shown in Fig.~\ref{fig:svanhovd_research_station}a. An ozone garden consists of selected plant species that are sensitive to ozone and are likely to display visible injuries. Cultivated species were, e.g. clover, tobacco, and potatoes. As shown in Fig.~\ref{fig:svanhovd_research_station}b, the plants had to be protected from herbivores with a wire-mesh fence. In 2018, we qualitatively observed visible ozone damage on semi-natural vegetation (clover) and crops (tobacco). In contrast, no such visible damage was found on the clovers in 2019, although it was observed on the sensitive tobacco cultivar. This indicates that the vegetation was more affected by ozone in 2018 than in 2019.
